Verification of concurrent design patterns with data

### 2019-02-14 2018-Arch-pNets/arch-pNets.tex
First pass ; shortening sec 3

\institute{%
  Inria Lille -- Nord Europe, Villeneuve d'Ascq, France\\
  \email{simon.bliudze@inria.fr}
  \and
	Univ Lyon, CNRS, ENS de Lyon, Inria, Universit\'e Claude Bernard Lyon 1, LIP, Lyon, France -- \email{ludovic.henrio@ens-lyon.fr}
  \and
    Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia-Antipolis, France\\
  \email{eric.madelaine@inria.fr}
}

@@

\begin{definition}[Open automaton]
	\label{defn:open-automaton}
	An \emph{open automaton} is a structure\\ $A =
	<J,\mathcal{S},s_0,\mathcal{T}>$ where:
	\begin{itemize}
		\item   $J$ is a  set of indices,
		\item   $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
		\item $\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist  $J'$ with  $J'
		\subseteq J$, such that $t$ is an open transition over  $J'$,
		and  $\mathcal{S}$.

	\end{itemize}
\end{definition}

@@

We build the semantics of open pNets as an open automaton where  the states are
tuples of states of the pLTSs at the leaves. We denote them $\triangleleft\ldots\triangleright$ for distinguishing tuple 
states from other tuples. given by 
Definition~\ref{def-states}. 
The open transitions first
 project the global state into states of the leaves, then apply
pLTS transitions on these states, and compose them with the sort of the holes according to the synchronisation defined in the synchronisation vectors. More precisely, each open transition between two states is defined as a composition of transition from some pLTSs and actions of some holes as follows: 1) the actions of the holes involved in the transition, 2) a guard  built from the synchronisation vectors coordinating the holes and the transitions involved; 3) assignments defined by the pLTSs transitions involved. The synchronisation vectors also define the global action. The global state evolution is also defined by the set of state transitions of the pLTSs involved in the pLTSs.

@@

\addSB{This section presents the extension of the theory of
  architectures \cite{AttieBBJS16-architectures-faoc} with data and
  briefly discusses a special case of priority models, called
  \emph{maximal progress}.}

@@

\subsubsection{Components and composition}
\label{secn:components}

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $\tuple{Q, q^0, V, \val{0}{}, P,
  \export, \goesto{}}$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    $\export : P \rightarrow 2^V$ 
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times (2^P\setminus \{\emptyset\}) \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports,
    %% $\emptyset \neq a \subseteq P$,
    a monotonic Boolean \emph{guard}
    %% $g \in \guards{V}$ 
    and a monotonic \emph{update assignment}.
    %% \hl{$u \in \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, only exported variables, \ie those in the set
  $\bigcup_{p \in P}\export(p)$, belong to the component interface and
  can be used for the interactions with other components (see
  \defn{im}).  In the context of this paper, we can simplify
  by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

@@

In this paper, we will use a running example, which is a refined
version of the Failure Monitor architecture
from~\cite{CubETH-case-study}.  Although \fig{schema:ArchFailure:BIP}
shows the full definition of this architecture, we will explain its
various elements progressively, as we define the corresponding
notions.

The model in \fig{schema:ArchFailure:BIP} has  components
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$(imer) and $\NameCtrl$(ontrol), with interfaces
$\btuple{
  \{t, z\},
  \{\PortTick, \PortCancel, \PortTOT\},
  \bigl\{\PortTick \mapsto \{t, z\}\bigr\}
}$,
and
$\btuple{
  \{\mathit{zone}\},
  \{\PortReset, \PortFail, \PortResume, \PortTOC\},
  \bigl\{\PortFail \mapsto \{\mathit{zone}\}\bigr\}
}$, respectively.
%
%% \[
%% \begin{array}{@{}c@{\hspace{3mm}}c@{}}
%%   \bigl(
%%   \{t, z\},
%%   \{\PortTick, \PortCancel, \PortTOT\},
%%   \export[\NameTimer]
%%   \bigr)\,,
%%   %% \qquad
%%   &
%%   \bigl(
%%   \{\mathit{zone}\},
%%   \{\PortReset, \PortFail, \PortResume, \PortTOC\},
%%   \export[\NameCtrl]
%%   \bigr)\,,
%%   \\
%%   \export[\NameTimer](p) =
%%   \begin{cases}
%%     \{t, z\}, & \text{if } p = \PortTick,\\
%%     \emptyset, & \text{otherwise,}
%%   \end{cases}
%%   %% \]
%%   %% and
%%   %% \[
%%   %% \quad
%%   &
%%   \export[\NameCtrl](p) =
%%   \begin{cases}
%%     \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
%%     \emptyset, & \text{otherwise.}
%%   \end{cases}
%% \end{array}
%% \]
%
%% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be an integer.  Variables $z$
and $\mathit{zone}$ are implicitly assumed to be integer intervals of
the form $[l,u]$.  For both types, we also implicitly assume the
natural ordering: increasing integers and interval inclusion.

@@

Transitions are labelled with ports of the
corresponding components, Boolean guards and update assignments on local
variables.  \Eg[,] the loop transition $\tstate{0}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{0}$ in the
component $\NameTimer$ is labeled by the port $\PortTick$.  It can be
fired only when the current value of the local variable $t$ is less
than $z.u$.
%% , where $z = [l, u]$ is a local variable representing an
%% interval
Upon firing, this transition increments the value of $t$ by $1$.  The
guards and update assignments of the transitions of $\NameCtrl$ are
omitted.  The default values are the constant predicate $\true$ (true) and
the empty assignment {\noop}, respectively.  
%\noteLH{It looks like Max is in zone, I do not understand}
%
Clearly, all guards and update assignments are monotonic.

@@

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of rule \eq{comp:semantics}
allows some of the variables to get new values before the
transition is actually fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise in \eq{comp:semantics} restricts the set of
variables, which can be affected by such data transfer, to those
that are exported through the ports participating in the
interaction.

@@

  An \emph{interaction model over $(V, P, \export)$} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \updates{V}$, such that,
  for any interaction $(a, g, u) \in \Gamma$, we have
  $g \in \guards{\export(a)}$ and $e \in \updates{\export(a)}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.
\end{definition}

%% \noteLH{supp should go outside the definition}
When speaking of a set of components or interfaces, we will always
assume that it satisfies the disjointness assumption of \defn{im}.
%
We call the \emph{support} of a set of ports $a \subseteq P$, denoted
$\supp{a}$, either the set $\setdef{i \in I}{a \cap P_i \neq
  \emptyset}$ (for $P = \bigcup_{i=1}^n P_i$) or the set $\setdef{B
  \in \cB}{a \cap P_B \neq \emptyset}$ (for $P = \bigcup_{B \in \cB}
P_B$).  The precise meaning of this notation will always be clear from
the context.  Intuitively, for an interaction $(a, g, u)$, $\supp{a}$
denotes the set of the participating components.

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = \tuple{Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{}}^{i \in I}$ with
  the interaction model $\Gamma$ over $\tuple{V, P, \export}$ (as in \defn{im})
  is the component $\Gamma(\cB) = \tuple{Q, q^0, V, \val{0}{}, P,
    \export, \goesto{}}$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather*}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather*}
  %% \todoSBin{Add the intuitive explanation}
\end{definition}

@@

Specifying interaction models as sets of sets of ports is not
practical due to their potentially exponential size.  An algebra of
connectors was introduced in \cite{BliSif08-acp-tc} in order to
structure interactions in BIP models.

Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of interactions, based on the
attributes of the nodes, which may be either \emph{trigger}
(triangles in \fig{schema:ArchFailure:BIP}) or \emph{synchron}
(bullets in \fig{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction is allowed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors
\emph{involving at least one of the triggers} is allowed.
%
The interaction model is defined as the set of all interactions
allowed by at least one of the connectors in the system.
%% The composition semantics of BIP systems consists in
%% firing exactly one interaction, enabled through at least one of the
%% top-level connectors, at each execution round.

For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} is a
two-level hierarchical connector.  In the second-level subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  Therefore, this
subconnector allows two interactions labelled $\{\PortFail\}$ and $\listset{\NameCtrl.\PortFail, \PortFail}$.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, whereas the second-level
subconnector is a synchron.  Hence, the overall connector allows the
interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%
  Such interactions are called \emph{singleton interactions}.
%
} and any interaction involving both {\NameTimer.\PortTick} and
\emph{some} interaction of the second-level subconnector.  Combining
the two levels we can see that the entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
\begin{gather*}
  \bigl(\{\NameTimer.\PortTick\},
  \true, \emptyset\bigr)
  %% \NameTimer.z := \NameTimer.z\bigr)
  ,
%
  \qquad
%
  \bigl(\{\PortFail, \NameTimer.\PortTick\},
  \true, \emptyset\bigr)
  %% \NameTimer.z := \NameTimer.z\bigr)
  ,
%
  \\
%
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \},
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
  .
\end{gather*}
%% \begin{align*}
%%   \bigl(\{\NameTimer.\PortTick\}&,
%%   \true, 
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
%%   \,.
%% \end{align*}

@@

\addSB{In addition to interaction models, BIP relies on the so-called
  \emph{priority models}, imposing a strict partial order on
  interactions.  Formal definitions are provided in \app{arch:appendix}.  Intuitively, an interaction can be fired only if all
  the higher-priority interactions available in the current state are
  disabled by their respective guards.
%
In the results and the running example of this paper, we will implicitly assume application of the so-called \emph{maximal
  progress} priority $\mu$, where $(a, g, u) \prec_\mu (b, h, w)$ iff  $a \subset b$ and $a \neq b$.  For instance, the
port $\NameTimer.\PortTick$ will never fire alone in a global state,
where the \noteSB{This turns out to be the first use of the word. I am tempted to pretend I did not notice.}\hl{dangling} port $\PortFail$ is also enabled.
In \secn{conclusion}, we discuss the possibility of generalising the
obtained results to other priority models.
}
%% Below, we will assume that the maximal progress priority $\mu$ is
%% applied after each composition with an interaction model.

@@

Architectures are partially defined BIP models, where \emph{dangling}
ports serve as placeholders for the eventual connection with
additional \emph{operand} components.

@@

An architecture $A$ enforces coordination constraints on the
components in $\cB$.  The interface $(V_A, P_A, \export[A])$ of an
architecture $A$ contains all ports of the coordinating
components $\cC$ and the dangling ports, which must belong to
the components in $\cB$.  In the application $A(\cB)$, the ports
belonging to $P_A$ can only participate in the interactions
defined by the interaction model $\Gamma$ of $A$.  Ports which do
not belong to $P_A$ are not restricted and can participate in any
interaction.  %% In particular, they can join the interactions in
%% $\Gamma$ (see \eq{im:extension}).  If the interface of the
%% architecture covers all ports of the system, \ie $P = P_A$, we
%% have $P\setminus P_A = \emptyset$ and the only interactions
%% allowed in $A(\cB)$ are those belonging to $\Gamma$.  Notice also
%% that the restrictions imposed by \defn{arch:application} on the
%% set of operand components $\cB$ ensure that an interaction in
%% $\IMextend{\Gamma}{P}$ can only refer to variables of the
%% participating components.
%
%% Finally, 
The definition of $\IMextend{\Gamma}{P}$ requires that
an interaction from $\Gamma$ be involved in every interaction
belonging to $\IMextend{\Gamma}{P}$.  To allow the ports from $P
\setminus P_A$ to be fired independently in $A(\cB)$, one must
have $(\emptyset, \true, \noop) \in \Gamma$.

@@

Below, for the sake of clarity, we start by presenting in
\secn{enc:nopri} the encoding of $A$ into pNets without accounting for
any priority model (see \secn{components}).  Then, in
\secn{enc:maxprog}, we slightly modify this encoding to include the
implicitly assumed maximal progress priority.

%****************************************************************
\subsection{Encoding without the priority model}
\label{secn:enc:nopri}

@@

\paragraph{Encoding the coordinators}
To encode a coordinator $C$, we 1)~introduce a fresh initial state with an additional
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~define a pNet action algebra that associates, as action parameter, an
additional Boolean value to each port, \add{3)~for each exported variable $x$, create the corresponding fresh variable $x'$ to allow updates during interactions,} 4)~create pNet transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 5)~create additional loop transitions marked by
$\false$.

>>>>

\institute{%
  Inria Lille -- Nord Europe, Villeneuve d'Ascq, France\\
  \email{simon.bliudze@inria.fr}
  \and
	Univ Lyon, EnsL, UCBL, CNRS, Inria,  LIP, F-69342, LYON Cedex 07, France\\ \email{ludovic.henrio@ens-lyon.fr}
  \and
    Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia-Antipolis, France\\
  \email{eric.madelaine@inria.fr}
}

@@

\begin{definition}[Open automaton]
	\label{defn:open-automaton}
	An \emph{open automaton} is a tuple $<\!\!J,\mathcal{S},s_0,\mathcal{T}\!\!>$ where:
%	\begin{itemize}
%		\item  
 $J$ is a  set of indices,
%		\item 
  $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
%		\item 
$\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist  $J'$ with  $J'
		\subseteq J$, such that $t$ is an open transition over  $J'$,
		and  $\mathcal{S}$.
%		
%	\end{itemize}
\end{definition}

@@

We build the semantics of open pNets as an open automaton where  the states are
tuples of states of the pLTSs at the leaves. We denote them $\triangleleft\ldots\triangleright$ for distinguishing tuple 
states from other tuples. 
%given by 
%Definition~\ref{def-states}. 
An open transition first
 project the global state into states of the leaves, then apply
pLTS transitions on these states, and compose them with the actions of the holes according to the synchronisation vectors. More precisely, each open transition between two states is defined as a composition of transition from some pLTSs and actions of some holes as follows: 1) the actions of the holes involved in the transition, 2) a guard  built from the synchronisation vectors coordinating the holes and the transitions involved; 3) assignments and global state change are defined by the pLTSs transitions involved. The synchronisation vectors also define the global action.

@@

This section presents the extension of the theory of
  architectures \cite{AttieBBJS16-architectures-faoc} with data and
  briefly discusses a special case of priority models, called
  \emph{maximal progress}.

@@

\subsubsection{Components and composition}
\label{secn:components}
\begin{definition}[Component]
  \label{defn:component}
  A component is a tuple $\tuple{Q, q^0, V, \val{0}{}, P,
  \export, \goesto{}}$ where\\[-4ex]
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports};
    $\export : P \rightarrow 2^V$ is
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times (2^P\setminus \{\emptyset\}) \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports,
    %% $\emptyset \neq a \subseteq P$,
    a monotonic Boolean \emph{guard}
    %% $g \in \guards{V}$ 
    and a monotonic \emph{update assignment}.
    %% \hl{$u \in \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{Only exported variables, belonging to a
  $\export(p)$, appear in the component interface  (see
  \defn{im}).  We omit here this separation between internal and exported variables.
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

@@

In this paper, we  use  a refined
version of the Failure Monitor architecture
from~\cite{CubETH-case-study} as a running example.  Although \fig{schema:ArchFailure:BIP}
shows the full definition of this architecture, we will explain its
various elements progressively.
\fig{schema:ArchFailure:BIP} shows  components
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$(imer) and $\NameCtrl$(ontrol), with interfaces
$\btuple{
  \{t, z\},
  \{\PortTick, \PortCancel, \PortTOT\},
  \bigl\{\!\PortTick \mapsto \{t, z\}\!\bigr\}
}$,
and
$\btuple{
  \{\mathit{zone}\},
  \{\PortReset, \PortFail, \PortResume, \PortTOC\},
  \bigl\{\!\PortFail \mapsto \{\mathit{zone}\}\!\bigr\}
}$ respectively.
%
%% \[
%% \begin{array}{@{}c@{\hspace{3mm}}c@{}}
%%   \bigl(
%%   \{t, z\},
%%   \{\PortTick, \PortCancel, \PortTOT\},
%%   \export[\NameTimer]
%%   \bigr)\,,
%%   %% \qquad
%%   &
%%   \bigl(
%%   \{\mathit{zone}\},
%%   \{\PortReset, \PortFail, \PortResume, \PortTOC\},
%%   \export[\NameCtrl]
%%   \bigr)\,,
%%   \\
%%   \export[\NameTimer](p) =
%%   \begin{cases}
%%     \{t, z\}, & \text{if } p = \PortTick,\\
%%     \emptyset, & \text{otherwise,}
%%   \end{cases}
%%   %% \]
%%   %% and
%%   %% \[
%%   %% \quad
%%   &
%%   \export[\NameCtrl](p) =
%%   \begin{cases}
%%     \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
%%     \emptyset, & \text{otherwise.}
%%   \end{cases}
%% \end{array}
%% \]
%
%% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be an integer.  Variables $z$
and $\mathit{zone}$ are  integer intervals of
the form $[l,u]$.  Both types are implicitly equipped with a
natural ordering: increasing integers and interval inclusion.

@@

Transitions are labelled with ports of the
corresponding components, Boolean guards and update assignments on local
variables.  \Eg[,] the loop transition $\tstate{0}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{0}$. % in the
%component $\NameTimer$ is labeled by the port $\PortTick$, the associated guarg and variable update. 
% It can be
%fired only when the current value of the local variable $t$ is less
%than $z.u$.
%%% , where $z = [l, u]$ is a local variable representing an
%%% interval
%Upon firing, this transition increments the value of $t$ by $1$.  
The
guards and update assignments of the transitions of $\NameCtrl$ are
omitted. By default, an omitted guard is  $\true$ (true) and an omitted assignment is empty {\noop}.  
%\noteLH{It looks like Max is in zone, I do not understand}
%
Clearly, all guards and update assignments are monotonic.

@@

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of rule \eq{comp:semantics}
allows some of the variables to get new values before the
transition is actually fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise in \eq{comp:semantics} restricts the set of
variables that can be affected by such data transfer to those
that are exported through the ports participating in the
interaction.

@@

  An \emph{interaction model over $(V, P, \export)$} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \updates{V}$, such that,
  for any interaction $(a, g, u) \in \Gamma$, we have
  $g \in \guards{\export(a)}$ and $u \in \updates{\export(a)}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.
\end{definition}

%% \noteLH{supp should go outside the definition}
We  
assume that all sets of components and interfaces satisfy the disjointness assumption above.
%
We call the \emph{support} of a set of ports $a \subseteq P$, denoted
$\supp{a}$, the set of the participating components. It is either the set $\setdef{i \in I}{a \cap P_i \neq
  \emptyset}$ (for $P = \bigcup_{i=1}^n P_i$) or the set $\setdef{B
  \in \cB}{a \cap P_B \neq \emptyset}$ (for $P = \bigcup_{B \in \cB}
P_B$).  The precise meaning of this notation will always be clear from
the context.

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = \tuple{Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{}}^{i \in I}$ with
  the interaction model $\Gamma$ over $\tuple{V, P, \export}$
  is the component $\Gamma(\cB) = \tuple{Q, q^0, V, \val{0}{}, P,
    \export, \goesto{}}$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather*}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather*}
  %% \todoSBin{Add the intuitive explanation}
\end{definition}

@@

Specifying interaction models as sets of sets of ports is not
practical due to their potentially exponential size.  An algebra of
connectors was introduced in \cite{BliSif08-acp-tc} in order to
structure interactions in BIP models.
Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of interactions, based on the
attributes of the nodes, which may be either \emph{trigger}
(triangles in \fig{schema:ArchFailure:BIP}) or \emph{synchron}
(bullets in \fig{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction is allowed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors
\emph{involving at least one of the triggers} is allowed.
%
The interaction model is defined as the set of all interactions
allowed by at least one of the connectors.
%% The composition semantics of BIP systems consists in
%% firing exactly one interaction, enabled through at least one of the
%% top-level connectors, at each execution round.

For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} of \fig{schema:ArchFailure:BIP} is a
two-level hierarchical connector.  In the  subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  This
subconnector allows two interactions: $\{\PortFail\}$ and $\listset{\NameCtrl.\PortFail, \PortFail}$.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, and the 
subconnector is a synchron. 
% Hence, the overall connector allows the
%interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%%
%  Such interactions are called \emph{singleton interactions}.
%%
%} and any interaction involving both {\NameTimer.\PortTick} and
%\emph{some} interaction of the second-level subconnector.  
%Combining
%the two levels we can see that t
The entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
%\begin{gather*}
$  \bigl(\{\NameTimer.\PortTick\},
  \true, \emptyset\bigr)
  %% \NameTimer.z := \NameTimer.z\bigr)
 $,
$  \bigl(\{\PortFail, \NameTimer.\PortTick\},
  \true, \emptyset\bigr)$
  %% \NameTimer.z := \NameTimer.z\bigr)
  ,
$
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \},
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
$
%  .
%\end{gather*}
%% \begin{align*}
%%   \bigl(\{\NameTimer.\PortTick\}&,
%%   \true, 
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
%%   \,.
%% \end{align*}

@@

\addSB{In addition to interaction models, BIP relies on 
  \emph{priority models} that impose  a strict partial order on
  interactions.  Formal definitions are provided in \app{arch:appendix}.  Intuitively, an interaction can be fired only if all
  the higher-priority interactions available in the current state are
  disabled by their respective guards.
%
In the next sections, we will implicitly assume application of the  \emph{maximal
  progress} priority $\mu$, where $(a, g, u) \prec_\mu (b, h, w)$ iff  $a \subset b$ and $a \neq b$.  For instance, the
port $\NameTimer.\PortTick$ will never fire alone if the port $\PortFail$ is also enabled.
In \secn{conclusion}, we discuss the generalisation of the
obtained results to other priority models.
}
%% Below, we will assume that the maximal progress priority $\mu$ is
%% applied after each composition with an interaction model.

@@

Architectures are partial BIP models, with \emph{dangling}
ports that serve as placeholders for the eventual connection with
 \emph{operand} components.

@@

An architecture $A$ enforces coordination constraints on the
components in $\cB$.  The interface $(V_A, P_A, \export[A])$ of an
architecture $A$ contains all ports of the coordinating
components $\cC$ and the dangling ports, which must belong to
the components in $\cB$.  In the application $A(\cB)$, the ports
belonging to $P_A$ can only participate in the interactions
defined by the interaction model $\Gamma$ of $A$.  Ports which do
not belong to $P_A$ are not restricted and can participate in any
interaction.  %% In particular, they can join the interactions in
%% $\Gamma$ (see \defn{arch:application}).  If the interface of the
%% architecture covers all ports of the system, \ie $P = P_A$, we
%% have $P\setminus P_A = \emptyset$ and the only interactions
%% allowed in $A(\cB)$ are those belonging to $\Gamma$.  Notice also
%% that the restrictions imposed by \defn{arch:application} on the
%% set of operand components $\cB$ ensure that an interaction in
%% $\IMextend{\Gamma}{P}$ can only refer to variables of the
%% participating components.
%
%% Finally, 
The definition of $\IMextend{\Gamma}{P}$ requires that
an interaction from $\Gamma$ be involved in every interaction
belonging to $\IMextend{\Gamma}{P}$.  To allow the ports from $P
\setminus P_A$ to be fired independently in $A(\cB)$, one must
have $(\emptyset, \true, \noop) \in \Gamma$.

@@

For the sake of clarity and conciseness, we define the encoding of $A$
into pNets without accounting for any priority model (see
\secn{components}).  Then, we provide a brief sketch of the
modification necessary to incorporate the implicitly assumed maximal
progress priority.

%% %****************************************************************
%% \subsection{Encoding without the priority model}
%% \label{secn:enc:nopri}

@@

\paragraph{Encoding the coordinators}
To encode a coordinator $C$, we 1)~introduce a fresh initial state with an additional
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~define a pNet action algebra that associates, as action parameter, an
additional Boolean value to each port, \add{3)~for each exported variable $x$, create the corresponding fresh variable $?x'$ to allow updates during interactions,} 4)~create pNet transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 5)~create additional loop transitions marked by
$\false$.
